---
title: "Covid Project"
author: "James Gabriel"
date: "`r Sys.Date()`"
output: html_document
---
## Covid Data Analysis

The dataset of our study contains daily & cumulative number of COVID-19 tests conducted, number of positive, hospitalized, recovered & death cases reported by country. In details here are the columns in the dataset:

1. Date: Date
1.  Continent_Name: Continent names
1. Two_Letter_Country_Code: Country codes
1. Country_Region: Country names
1. Province_State: States/province names; value is All States when state/provincial level data is not available
1. positive: Cumulative number of positive cases reported.
1. active: Number of active cases on that day.
1. hospitalized: Cumulative number of hospitalized cases reported.
1. hospitalizedCurr: Number of actively hospitalized cases on that day.
1. recovered: Cumulative number of recovered cases reported.
1. death: Cumulative number of deaths reported.
1. total_tested: Cumulative number of tests conducted.
1. daily_tested: Number of tests conducted on the day; if daily data is unavailable, daily tested is averaged across number of days in between.
1. daily_positive: Number of positive cases reported on the day; if daily data is unavailable, daily positive is averaged across number of days in.

This analysis tries to provide an answer to this question: Which countries have had the highest number of positive cases against the number of tests?

Firstly, we'd load in the `tidyverse` package as we would be using this for our analysis.
```{r}
library(tidyverse)
```
```{r}
covid_df <- read_csv('covid.csv')
```
To have a brief idea of what the dataset looks like.

```{r}
dim(covid_df)
```
Displaying the column names.

```{r}
vector_cols <- colnames(covid_df)
```
```{r}
vector_cols
```
```{r}
class(covid_df$Country_Region)
```
```{r}
class(vector_cols)
```
Let's diaplay a few rows of the dataset to explore it visually.

```{r}
head(covid_df)
```
The `glimpse` function is used to have an idea of the different data types of each column with a few samples.

```{r}
glimpse(covid_df)
```

Considering most of the data is either referring to a particular province or the entire country, it is necessary to have consistency to ensure proper analysis. For this reason, I'm filtering out the data to contain only rows where the `Province_State` says `All States`. 
I would also be dropping the `Province_State` column.

```{r}
covid_df_all_states <- covid_df %>% filter(Province_State=='All States') %>% select(-Province_State)
```
```{r}
covid_df_all_states
```

The data also contains daily and cummulative aggregation. To ensure consistency, I would be filtering for only columns that record daily changes.

```{r}
covid_df_all_states_daily <- covid_df_all_states %>% select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)
```
```{r}
covid_df_all_states_daily
```

I then checked to see the total number of `tested`, `positive`, `active`, and `hospitalized` cases for each country present in the dataset. After this, I sorted the result in descending order and filtered to just the top 10 countries with the highest `tested` cases.

```{r}
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% 
                                group_by(Country_Region) %>% 
                                summarise(
                                  tested = sum(daily_tested),
                                  positive = sum(daily_positive),
                                  active = sum(active),
                                  hospitalized = sum(hospitalizedCurr)) %>% 
                                arrange(-tested)
```

```{r}
covid_df_all_states_daily_sum
```

```{r}
covid_top_10 <- head(covid_df_all_states_daily_sum, 10)
```
```{r}
covid_top_10
```
Assigning each of the columns in the top 10 result to various vectors to use for further analysis.


```{r}
countries <- covid_top_10 %>% pull(Country_Region)

tested_cases <- covid_top_10 %>% pull(tested)

positive_cases <- covid_top_10 %>% pull(positive)

active_cases <- covid_top_10 %>% pull(active)

hospitalized_cases <- covid_top_10 %>% pull(hospitalized)
```

```{r}
tested_cases
```

```{r}
names(tested_cases)
```
Giving names to the values in the vector.

```{r}
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
```

```{r}
names(active_cases)
```
```{r}
active_cases
```

To answer the question asked earlier, feature engineering had to be put in place to generate a new result that calculates the total number of positive cases divided by the tested cases to find the countries with the highest ratio.

```{r}
positive_cases/tested_cases
```

`United Kingdom`, `United States`, and `Turkey` took the lead with the highest ratios.

```{r}
positive_tested_top_3 <- positive_cases/tested_cases
```
```{r}
positive_tested_top_3 <- positive_tested_top_3[c(7,1,5)]
```
```{r}
positive_tested_top_3
```

A matrix was created to store this information.

```{r}
united_kingdom <- c(0.11, 1473672, 166909, 0, 0)
united_states <- c(0.10, 17282363, 1877179, 0, 0)
turkey <- c(0.08, 2031192, 163941, 2980960, 0)
```
```{r}
covid_mat <- rbind(united_kingdom, united_states, turkey)
```
```{r}
colnames(covid_mat) <- c("Ratio", "tested", "positive", "active", "hospitalized")
```
```{r}
covid_mat
```
```{r}
question <- "Which countries have had the highest number of positive cases against the number of tests?"
```

```{r}
answer <- c("Positive tested cases" = positive_tested_top_3)
```

```{r}
data_structure_list <- list( c(covid_df, covid_df_all_states, covid_df_all_states_daily, covid_top_10),
                             covid_mat, c(vector_cols, countries))
```

```{r}
covid_analysis_list <- list(question, answer, data_structure_list)
```

```{r}
covid_analysis_list[[2]]
```

## Conclusion

From the analysis, it was observed that the country with the highest number of positive cases against the number of tested cases was the `United Kingdom`, followed closely by the `United States` and then `Turkey`.